Path Analysis Data Development - Chapter 2

### 2016-03-22 sem-data-development.Rmd
Add infection rule.
I'm not yet confident in this calculation.

```{r join umca and qusp data}
umca_plots
umca_plots$sample_year <- as.integer(umca_plots$sample_year)
plot_qusp # right now this is broken down by oak species
qusp4_plots <- stems %>%
      select(plotid, date, species, canker, year, stem_status) %>%
      filter(species == "quag" | species == "quke" | species == "quch" |
                   species == "quwi", stem_status == "Alive") %>% 
      group_by(plotid, year, species) %>% 
      summarise(uninfected_oak_ct = length(which(canker==0)), 
                infected_oak_ct = length(which(canker > 0)), 
                tot_oak = length(species))

@@

oak_sod <- qusp4_plots %>% group_by(plotid, year) %>%
      summarise(uninf_oak_ct = sum(uninfected_oak_ct), 
                inf_oak_ct = sum(infected_oak_ct), 
                tot_oak_ct = sum(c(uninfected_oak_ct, infected_oak_ct)))
oak_sod
oak_sod <- rename(oak_sod, sample_year = year)
# oak_sod$plotid <- tolower(oak_sod$plotid)

@@

hrsblw14_wet_lag <- as.tbl(select(wet_hrs, plotid, sample_year, hrsblw14_wet))
avgtmax_wet_lag <- as.tbl(select(wet_hrs, plotid, sample_year, avgtmax_wet))
hrs1422_wet_lag <- as.tbl(select(wet_hrs, plotid, sample_year, hrs1422_wet))
wet_hrs_lag <- select(wet_hrs_lag, plotid, sample_year)
wet_hrs_lag <- wet_hrs_lag %>% 
      mutate(sample_year_minus1 = sample_year - 1,
             sample_year_minus2 = sample_year - 2,
             sample_year_minus3 = sample_year - 3)

@@

# remove unnecessary data
rm(list = ls(pattern = "^rain_rs"))
rm(list = ls(pattern = "^rain_ss"))
rm(list = ls(pattern = "^rain2d"))
rm(list = ls(pattern = "*ds_lag"))
rm(list = ls(pattern = "*0_rs_lag")); rm(avgtmax_rs_lag)
rm(list = ls(pattern = "*wet_lag"))

>>>>

```{r join umca and qusp data}
umca_plots
umca_plots$sample_year <- as.integer(umca_plots$sample_year)
plot_qusp # right now this is broken down by oak species

summary(stems)
# calculate the number of times a canker was observed on each stem
cankers_observed <- stems %>%
      select(plotid, year, species, tag, canker, stem_status) %>% 
      filter(species == "quag" | species == "quke" | species == "quch" |
                   species == "quwi", stem_status == "Alive") %>% 
      group_by(plotid, species, tag) %>% 
      summarise(times_canker_observed = sum(canker, na.rm = T))
summary(cankers_observed)
cankers_observed <- droplevels(cankers_observed)

## If canker was observed at least twice for a stem then "infection confirmed"
## This is Sarah's "twice cankered always infected rule" - more conservative
cankers_observed$infection_confirmed_2plus <- ifelse(
      cankers_observed$times_canker_observed >= 2, 1, 0
)
## If canker was observed just once then "infection confirmed"
## This is a "once cankered always infected rule" - less conservative
cankers_observed$infection_confirmed_1plus <- ifelse(
      cankers_observed$times_canker_observed >= 1, 1, 0
)
## Both of these approaches are an effort to better represent oak infection, instead of just based on whether or not a canker was observed for a given year.

stems <- left_join(stems, cankers_observed, by = c("plotid", "species", "tag"))

stems <- stems %>% mutate(
      inf_2plus = ifelse(
            year == 2003 & canker == 1 & infection_confirmed_2plus == 1, 1, 0)
) %>% mutate(inf_2plus = ifelse(
      year == 2004 & canker == 1 & infection_confirmed_2plus == 1 & inf_2plus != 1, 1, 0)) %>% 
      mutate(inf_2plus = ifelse(
            year == 2005 & canker == 1 & infection_confirmed_2plus == 1 & inf_2plus != 1, 1, 0)) %>% 
      mutate(inf_2plus = ifelse(
            year == 2006 & canker == 1 & infection_confirmed_2plus == 1 & inf_2plus != 1, 1, 0)) %>% 
      mutate(inf_2plus = ifelse(
            year == 2007 & canker == 1 & infection_confirmed_2plus == 1 & inf_2plus != 1, 1, 0)) %>% 
      mutate(inf_2plus = ifelse(
            year == 2008 & canker == 1 & infection_confirmed_2plus == 1 & inf_2plus != 1, 1, 0)) %>% 
      mutate(inf_2plus = ifelse(
            year == 2009 & canker == 1 & infection_confirmed_2plus == 1 & inf_2plus != 1, 1, 0)) %>%
      mutate(inf_2plus = ifelse(
            year == 2010 & canker == 1 & infection_confirmed_2plus == 1 & inf_2plus != 1, 1, 0)) %>% 
      mutate(inf_2plus = ifelse(
            year == 2011 & canker == 1 & infection_confirmed_2plus == 1 & inf_2plus != 1, 1, 0)) %>% 
      mutate(inf_2plus = ifelse(
            year == 2012 & canker == 1 & infection_confirmed_2plus == 1 & inf_2plus != 1, 1, 0)) %>% 
      mutate(inf_2plus = ifelse(
            year == 2014 & canker == 1 & infection_confirmed_2plus == 1 & inf_2plus != 1, 1, 0))

qusp_stems <- stems %>% 
      select(plotid, year, tag, species, canker, stem_status, infection_confirmed_1plus, infection_confirmed_2plus) %>% 
      filter(species == "quag" | species == "quke" | species == "quch" |
                   species == "quwi", stem_status == "Alive")

qusp4_plots <- stems %>%
      select(plotid, species, canker, year, stem_status, infection_confirmed_1plus, infection_confirmed_2plus) %>%
      filter(species == "quag" | species == "quke" | species == "quch" |
                   species == "quwi", stem_status == "Alive") %>% 
      group_by(plotid, year, species) %>% 
      summarise(uninfected_oak_ct = length(which(canker == 0)), 
                infected_oak_ct = length(which(canker > 0)), 
                tot_oak = length(species),
                inf_oak_1plus_ct = length(which(infection_confirmed_1plus > 0)),
                inf_oak_2plus_ct = length(which(infection_confirmed_2plus > 0)))
summary(qusp4_plots)
# write.csv(qusp4_plots, "analysis/data/qusp4_infections.csv", row.names = F)
qusp4_plots <- read.csv("analysis/data/qusp4_infections.csv")

@@

oak_sod <- qusp4_plots %>% group_by(plotid, year) %>%
      summarise(uninf_oak_ct = sum(uninfected_oak_ct), 
                inf_oak_ct = sum(infected_oak_ct), 
                tot_oak_ct = sum(c(uninfected_oak_ct, infected_oak_ct)),
                inf_oak_1plus_ct = sum(inf_oak_1plus_ct),
                inf_oak_2plus_ct = sum(inf_oak_2plus_ct))
oak_sod
oak_sod <- rename(oak_sod, sample_year = year)
summary(oak_sod)
# oak_sod$plotid <- tolower(oak_sod$plotid)

@@

hrsblw14_wet_lag <- as.tbl(select(wet_hrs, plotid, sample_year, hrsblw14_wet))
avgtmax_wet_lag <- as.tbl(select(wet_hrs, plotid, sample_year, avgtmax_wet))
hrs1422_wet_lag <- as.tbl(select(wet_hrs, plotid, sample_year, hrs1422_wet))
wet_hrs_lag <- select(wet_hrs, plotid, sample_year)
wet_hrs_lag <- wet_hrs_lag %>% 
      mutate(sample_year_minus1 = sample_year - 1,
             sample_year_minus2 = sample_year - 2,
             sample_year_minus3 = sample_year - 3)

@@

# remove unnecessary data
rm(list = ls(pattern = "^rain_rs"))
rm(list = ls(pattern = "^rain_ss"))
rm(list = ls(pattern = "^rain2d"))
rm(list = ls(pattern = "*ds_lag"))
rm(list = ls(pattern = "*0_rs_lag")); rm(avgtmax_rs_lag)
rm(list = ls(pattern = "*wet_lag")); #rm(wet_hrs_lag1)
